---
title: "Kaggle: Rossmann Store Sales"
author: "Liudmila Dzemiantsova"
date: "October 16, 2015"
output: html_document
---

**Description**

The Website of the Kaggle competition is https://www.kaggle.com/c/rossmann-store-sales.
Rossmann operates over 3,000 drug stores in 7 European countries. Currently, Rossmann store managers are tasked with predicting their daily sales for up to six weeks in advance. Store sales are influenced by many factors, including promotions, competition, school and state holidays, seasonality, and locality. With thousands of individual managers predicting sales based on their unique circumstances, the accuracy of results can be quite varied.

**Task**

In their first Kaggle competition, Rossmann is challenging you to predict 6 weeks of daily sales for 1,115 stores located across Germany. Reliable sales forecasts enable store managers to create effective staff schedules that increase productivity and motivation. 

**Programming and Software**

The code is written with R v.3.2.1 using RStudio v.0.99.484

**Libraries**

```{r}
source("multiplot.R")
```

```{r, echo=FALSE}
library(GGally)
library(doParallel)
library(caTools)
library(pROC)
library(caret)
library(readr)
library(mice)
library(corrplot)
library(maptools)
library(maps)
library(reshape2)
```

**Read The Training Data**

```{r}
registerDoParallel(cores=1)
training <-  read.csv("train.csv")
dim(training)
sum(is.na(training))
md.pattern(training)
```

**Exploratory Analysis**

```{r}
fun <- function(x){ifelse(x == 0, "close", "open") }
training$Open <- as.factor(sapply(training$Open, fun))
table(training$Open, training$DayOfWeek)

df1 <- data.frame(table(training$Open))
colnames(df1) <- c('','Freq')
df1$Perc <- df1$Freq / sum(df1$Freq) * 100
df1

fun <- function(x){ifelse(x == 0, "no", "yes") }
training$SchoolHoliday <- as.factor(sapply(training$SchoolHoliday, fun))
table(training$SchoolHoliday, training$DayOfWeek)

SchHoli.no <- vector()
SchHoli.yes <- vector()
for(i in 1:7){
   SchHoli.no[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$SchoolHoliday == "no",]$Sales)
   SchHoli.yes[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$SchoolHoliday == "yes",]$Sales)
}
DayOfWeek <- c(1, 2, 3, 4, 5, 6, 7)
df1 <- data.frame(DayOfWeek = DayOfWeek, Day = SchHoli.no, Holiday = SchHoli.yes)
df2 <- melt(data = df1, id.vars = "DayOfWeek")
ggplot(data = df2, aes(x = DayOfWeek, y = value, colour = variable)) +
       geom_point(size = 5)+
       ylab("Sales") + xlab("Day Of Week")+
       ggtitle("Sales Per School Holidays")

StHoli.no <- vector()
StHoli.a <- vector()
StHoli.b <- vector()
StHoli.c <- vector()
for(i in 1:7){
   StHoli.no[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$StateHoliday == "0",]$Sales)
   StHoli.a[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$StateHoliday == "a",]$Sales)
   StHoli.b[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$StateHoliday == "b",]$Sales)
   StHoli.c[i] <- sum(training[training$DayOfWeek==i & training$Open == "open" & training$StateHoliday == "c",]$Sales)
}
df1 <- data.frame(DayOfWeek = DayOfWeek, Holiday.A = StHoli.a, Holiday.B = StHoli.b, Holiday.C = StHoli.c, Day = StHoli.no)
df2 <- melt(data = df1, id.vars = "DayOfWeek")
ggplot(data = df2, aes(x = DayOfWeek, y = value, colour = variable))+
       geom_point(size = 5)+
       ylab("Sales") + xlab("Day Of Week")+
       ggtitle("Sales Per State Holidays")

df1 <- data.frame(DayOfWeek = DayOfWeek, Holiday.A = StHoli.a, Holiday.B = StHoli.b, Holiday.C = StHoli.c)
df2 <- melt(data = df1, id.vars = "DayOfWeek")
ggplot(data = df2, aes(x = DayOfWeek, y = value, colour = variable))+
       geom_point(size = 5)+
       ylab("Sales") + xlab("Day Of Week")+
       ggtitle("Sales Per State Holidays")

training$Date <- as.character(training$Date)
fun1 <- function(x){strsplit(x, "-")[[1]][[1]]}
training$Year <- as.factor(sapply(training$Date, fun1))
fun2 <- function(x){strsplit(x, "-")[[1]][[2]]}
training$Month <- as.factor(sapply(training$Date, fun2))
fun3 <- function(x){strsplit(x, "-")[[1]][[3]]}
training$Day <- as.factor(sapply(training$Date, fun3))

Sales.2015 <- sum(training[training$Year=="2015" & training$Open == "open",]$Sales)
Sales.2014 <- sum(as.numeric(training[training$Year=="2014" & training$Open == "open",]$Sales))
Sales.2013 <- sum(as.numeric(training[training$Year=="2013" & training$Open == "open",]$Sales))

Year <- c(2013, 2014, 2015)
Sales <- c(Sales.2013, Sales.2014, Sales.2015)
df1 <- data.frame(Year = as.integer(Year), Sales = Sales)
ggplot(data = df1, aes(x = Year, y = Sales))+
       geom_point(size = 5)+
       ylab("Sales") + xlab("Year")+
       ggtitle("Sales Per Year")

df.2015 <- training[training$Year=="2015" & training$Open == "open",]
df.2014 <- training[training$Year=="2014" & training$Open == "open",]
df.2013 <- training[training$Year=="2013" & training$Open == "open",]
Month.2015 <- vector()
Month.2014 <- vector()
Month.2013 <- vector()
for(i in 1:12){
   Month.2015[i] <- sum(df.2015[as.numeric(df.2015$Month) == i,]$Sales)
   Month.2014[i] <- sum(df.2014[as.numeric(df.2014$Month) == i,]$Sales)
   Month.2013[i] <- sum(df.2013[as.numeric(df.2013$Month) == i,]$Sales)
}
Month <- c(1,2,3,4,5,6,7,8,9,10,11,12)
df1 <- data.frame(Month = Month, Year_2015 = Month.2015, Year_2014 = Month.2014, Year_2013 = Month.2013)
df2 <- melt(data = df1, id.vars = "Month")
ggplot(data = df2, aes(x = Month, y = value, colour = variable)) +
       geom_point(size = 5)+
       ylab("Sales") + xlab("Month")+
       ggtitle("Sales Per Month")

training <- training[,-3]
ggpairs(training[training$Open == "open",], columns = c(1,3,4), legends=TRUE, lower = list(continuous = "points"),
        diag = list(continuous = "density"), axisLabels = "show", colour = "Year",
        title = "Density Distribution Of Store, Sales And Custemers With Respect To Year") # uper = "blank"

ggpairs(training[training$Open == "open",], columns = c(1,3,4), legends=TRUE, lower = list(continuous = "points"),
        diag = list(continuous = "density"), axisLabels = "show", colour = "Month",
        title = "Density Distribution Of Store, Sales And Custemers With Respect To Month") # uper = "blank"

training_numr = training[, sapply(training, is.numeric)]
corrplot(cor(training_numr, method = "pearson"), method = "number",order = "hclust",type='lower', diag=F,addCoefasPercent=T)
```